Variable Density Tesseroids: Gravity fields calculation in spherical coordinates using variable densities in depth

### 2017-07-31 manuscript/manuscript.tex
Minor corrections in equations
Minor modification in manuscript
Modified exponential function in manuscript: now it's a decreasing function.
Minor correction in exponential function expression
Continue writing implementation

\begin{equation}
    \cos\psi = \sin\phi', \quad
    \ell = \sqrt{r'^2 + 2 r' r \sin\phi' + r^2}.
\end{equation}

@@

\begin{equation}
    V_\text{sh}(r) = 2\pi G 
    \int\limits_{-\frac{\pi}{2}}^\frac{\pi}{2}
    \int\limits_{R_1}^{R_2}
    \frac{\rho(r') {r'}^2 \cos\phi'}{\sqrt{r'^2 + 2 r' r \sin\phi' + r^2}}
    \, dr' d\phi',
\end{equation}

@@

\begin{equation}
    \rho(r') = Ae^\frac{r'-\Delta h}{b},
\end{equation}

\noindent we have used \citet{sagemath} to obtain the following expression for the potential:

\iftwocol{
\begin{equation}
    \begin{split}
        V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
        Ab \, e^{-\frac{\Delta h}{b}} \Big[ 
        & (R_2^2 - 2R_2 b + 2b^2)e^\frac{R_2}{b} - \\
        & (R_1^2 - 2R_1 b + 2b^2)e^\frac{R_1}{b} \Big].
    \end{split}
\end{equation}
}{
\begin{equation}
    V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
    Ab \, e^{-\frac{\Delta h}{b}} \left[ 
    (R_2^2 - 2R_2 b + 2b^2)e^\frac{R_2}{b} - 
    (R_1^2 - 2R_1 b + 2b^2)e^\frac{R_1}{b} \right].
\end{equation}
}

@@

We have developed a Python library that computes the gravity fields generated by any Tesseroid with variable density in depth that is freely available under the BSD 3-clause open-source library and can be downloaded from the following repository: \todo{agregar repo, vale la pena poner el link?}

@@

>>>>

\begin{equation}
    \cos\psi = \sin\phi', \quad
    \ell = \sqrt{r'^2 + 2 r r' \sin\phi' + r^2}.
\end{equation}

@@

\begin{equation}
    V_\text{sh}(r) = 2\pi G 
    \int\limits_{-\frac{\pi}{2}}^\frac{\pi}{2}
    \int\limits_{R_1}^{R_2}
    \frac{\rho(r') {r'}^2 \cos\phi'}{\sqrt{r'^2 + 2 r r' \sin\phi' + r^2}}
    \, dr' d\phi',
\end{equation}

@@

\begin{equation}
    \rho(r') = Ae^{-(r' - \Delta h)/b},
\end{equation}

\noindent the potential can be written as follows:

\iftwocol{
\begin{equation}
    \begin{split}
        V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
        Ab \, e^\frac{\Delta h}{b}
        \Big[
        & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
        & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
        \Big].
    \end{split}
\label{eq:shell-pot-exp}
\end{equation}
}{
\begin{equation}
    V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
    Ab \, e^\frac{\Delta h}{b}
    \Big[
    & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
    & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
    \Big].
\label{eq:shell-pot-exp}
\end{equation}
}

@@

We have developed a Python library that computes the gravity fields generated by any Tesseroid with variable density in depth through the GLQ to numerically approximate the volume integrals. It's freely available under the BSD 3-clause open-source library and can be downloaded from the following repository: \todo{agregar repo, vale la pena poner el link?}

@@

From equation \ref{eq:glq-var-dens} we saw that the GLQ approximates the volume integrals as the gravitational effect of $N_\lambda N_\phi N_r$ point masses, where $N_i$ are the orders of the quadrature for the integral on each coordinate, with $i \in \{ \lambda, \phi, r \}$.

If we set the GLQ orders to fixed values, it's easy to conclude that the nearer the computation point is to the tesseroid, the accuracy of the approximation is reduced and the point masses effect becomes more noticeable than the tesseroid we want to approximate.
A more complete analysis of this conclusion can be found in \citet{Ku1977, Li2011, Uieda2016}.

Instead of increasing the GLQ orders, what leads to a more time consuming computation, adaptative discreatization algorithms have been applied in previous works \citep{Li2011, Uieda2016}.
They basically consist in dividing the tesseroid based on a ratio between its size and the distance between its geometric centre and the computation point.

In order to make the computation more efficient, we use only second order GLQ for each integration, and we ensure the accuracy of the method through the application of the modified adaptative discretization algorithm developed by \citet{Uieda2016}.

Given an arbitrary tesseroid, \citet{Uieda2016} redefined its dimensions: $L_\lambda$, $L_\phi$ and $L_r$, where the former ones are the arc-distances measured along its top surface and $L_r$ is the difference between the radii of the top and bottom surfaces.
We must check that the following inequality is satisfied:

\begin{equation}
\frac{d}{L_i} \geq D,
\label{eq:distance-size-ratio}
\end{equation}

\noindent where $i \in \{\lambda, \phi, r\}$, $d$ is the distance between the geometric centre of the tesseroid and the computation point and $D$ is a predefined value called distance-size ratio.
If the inequality \ref{eq:distance-size-ratio} is not held for any spherical coordinate, then the tesseroid must be divided by half in that direction.

In this way, the numerical approximation efficiently improves its accuracy and
gives the user more control over the computation by allowing to stop the division of the tesseroids when their dimensions are lower than a certain value (e.g. less than 10cm).

On the other hand, the accuracy is exclusively controlled by the value assigned to the distance-size ratio $D$, which cannot be easily related to the error of the approximation.
In order to solve this, \citet{Uieda2016} compared the numerical computation with the analytical solution for a spherical shell with homogeneous density.

We have performed similar tests, but now with variable density spherical shell. From the infinite density functions that can be proposed, we have chosen two characteristic ones: a linear and an exponential functions.
The first one represents the simpler variation, while the second one symbolise the fastest varying function.
The analytical solutions for these two types of functions were previously obtained and shown in equations \ref{eq:shell-pot-linear} and \ref{eq:shell-pot-exp}.
